---
title: "Classification when all Independent variables are Categorical"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading in the Libraries

install.packages("testthat")
install.packages("gridExtra")
install.packages("corrplot")
install.packages("GGally")
install.packages("e1071")
install.packages("lsr")
install.packages("logistf")
install.packages("caTools")
install.packages("randomForest")
install.packages("rpart")
install.packages("rpart.utils")
install.packages("rpart.plot")
install.packages("rockchalk")
install.packages("brglm")
install.packages("plotrix")

```{r cars}
library(data.table)
library(testthat)
library(gridExtra)
library(corrplot)
library(GGally)
library(ggplot2)
library(e1071)
library(dplyr)
library(lsr)
library(logistf)
library(caret)
library(caTools)
library(randomForest)
library(rpart)
library(rpart.utils)
library(rpart.plot)
library(rockchalk)
library(brglm)
library(plotrix)
```

#### Exploratory Data Analysis

Since all the variables are cetegorical.. chi square test tells the association between all categorical variable loop on all variables in dataframe for chi square test

```{r read}
mushrooms = read.csv("C:/Users/this pc/Documents/Kaggle/mushrooms.csv",header=T)

mushrooms$freq=1

mat = matrix(nrow=23,ncol=23, byrow = TRUE, dimnames= list(names(mushrooms),names(mushrooms)))

for(i in 1:ncol(mushrooms))
{
  for(j in 1:ncol(mushrooms))
  {
    print(i)
    print(j)
    temp = xtabs(freq~mushrooms[,j]+mushrooms[,i], data = mushrooms)
    ##temp2= chisq.test(temp)
    mat[i,j]= cramersV(temp)
  }
}

```


#### Now plotting a matrix using 2dMatplot

``` {r eplot}
cellcol<-color.scale(cbind(mat,c(-1,rep(1,22))),c(0,1),0,c(1,0))[,1:13]
par(mar = c(8,8,0.1,0.1) + 0.3)

color2D.matplot(mat,cellcolors=cellcol,show.legend=TRUE,show.values=1,
                axes=FALSE, xlab="",ylab="")

axis(1,at=0.5:22.5,las=2,labels=colnames(mat))
axis(2,at=0.5:22.5,las=2,labels=rev(rownames(mat)))
```