---
title: "HR Analytics - Leaver Prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


#### Loading the libraries and the dataset


``` {r s1, warning = FALSE}
library(dplyr)
library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(e1071)

```

``` {r s2,echo=FALSE,}
h = read.csv("C:/Users/this pc/Documents/Kaggle/HRA/HR_comma_sep.csv",header = TRUE)

```

#### Exploratory Analysis is performed to understand these questions


* (general thoughts marked in brackets) *
* (the input levels - manhours,time spend) *
* (output levels - projects, evaluation) *
* (reaction (how do employees perceive/infer/reflect) - satisfaction levels) *
* (action (wat is the outcomes/company perceive) - promotion levels, salary) *
* (issues - disciplinary, accidents, etc) *

#### what kind of people are leaving?

``` {r s3}
h %>% filter(left == 1) ->h.left
summary(h.left$satisfaction_level)


ggplot(h.left, aes(x=satisfaction_level)) + geom_histogram(binwidth=0.10)
ggplot(h.left, aes(x=last_evaluation)) + geom_histogram(binwidth=0.10)
ggplot(h.left, aes(x=number_project,y=..prop..)) + geom_bar()
ggplot(h.left, aes(x=time_spend_company)) + geom_histogram(binwidth=1)
ggplot(h.left, aes(x=salary,y=..count..)) + geom_bar()
ggplot(h.left, aes(x=promotion_last_5years,y=..count..)) + geom_bar()

```

nearly 900 have very low satification levels and 1500 have less than average satisfaction levels  
for nearly 1500 of them evaluation scores where less than 0.5  
nearly 45% of them had only 2 projects in hand  
tenure is between 3 to 5 years  
nearly 2000 of them out of 3500 had low salary  
nearly all of them have not been promoted in the last 5 years  

#### are there any good talents leaving the company?
even though some of them have less evaluation scores and satisfaction levels  
some of them to be good talent and have had a good tenure of 3 to 5 years  
one of the reasons seems to be : no promotion and low pay  

#### how many left due to work_accident?

``` {r s5}
h.left %>% group_by(Work_accident) %>% summarize( count = n(),percent = round(n()/nrow(h.left)*100))
```

only 169 or 5% of them had left due to accidents

### is there any danger of further attrition of good talent ?

``` {r s6}
h %>% filter(left==0) %>% 
      filter(last_evaluation >0.7|satisfaction_level>0.7|number_project>3)->h.goodnotleft

ggplot(h.goodnotleft, aes(x=salary,y=..count..)) + geom_bar()
ggplot(h.goodnotleft, aes(x=promotion_last_5years,y=..count..)) + geom_bar()
```

a big yes - nearly 4500 have been paid low. out of 10002, almost none of them have not been promoted

``` {r s7}
ggplot(h.goodnotleft, aes(x=number_project,y=..prop..)) + geom_bar()
ggplot(h.goodnotleft, aes(x=average_montly_hours)) + geom_histogram(binwidth = 50)
```

they seem to be overworked with more than 50 hours per week for 6000 of them  
65% are handling 4 or more projects  


#### we create model to retain the most valuable employees

``` {r s8}
h %>% filter(last_evaluation >0.7|satisfaction_level>0.7|number_project>3)->h.good

h.good$left <- as.factor(h.good$left)
table(h.good$left)

train_control<- trainControl(method="cv", number=5, repeats=3)
# train the model 
rpartmodel<- train(left~., data=h.good, trControl=train_control, method="rpart")
# make predictions
predictions<- predict(rpartmodel,h.good)
hr_model_tree<- cbind(h.good,predictions)
# summarize results
confusionMatrix<- confusionMatrix(hr_model_tree$predictions,hr_model_tree$left)
confusionMatrix
```

``` {r s9}
e1071model2 <- train(left~., data=h.good, trControl=train_control, method="nb")
# make predictions
predictions<- predict(e1071model2,h.good)
e1071modelbinded <- cbind(h.good,predictions)
# summarize results
confusionMatrix<- confusionMatrix(e1071modelbinded$predictions,e1071modelbinded$left)
confusionMatrix
```